Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Istanbul version \YellowPaperVersionNumber}}

### 2021-06-22 Paper.tex
Revert an unintended change from PR #805
Introduce protocol update block numbers (#804)
* Introduce fork block numbers

* Don't use fork for protocol updates

Sometimes, a path follows a new protocol from a particular height (block number).
This document describes one version of the protocol, namely the \textit{Istanbul} version defined in EIP-1679 by \cite{EIP-1679}.
In order to follow back the history of a path, one must reference multiple versions of this document.

Occasionally actors cannot agree on a protocol change, and a permanent fork occurs.
In order to distinguish between diverged blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
For the Ethereum main network
\begin{equation}
  \linkdest{chain_id}
  \beta = 1
\end{equation}

@@

&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \mathbf{b} \rVert$ \\

@@

@@

0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current\\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\

@@

\begin{tabu}{\usetabu{opcodes}}
\midrule
0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
\midrule
\linkdest{extcodehash}{}0x3f & {\small EXTCODEHASH} & 1 & 1 & Get hash of an account's code. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
\begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}) \\ \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}} & \text{otherwise} \end{cases}$ \\
\bottomrule
\end{tabu}

@@

G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, t) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\

>>>>

Sometimes, a path follows a new protocol from a particular height (block number).
This document describes one version of the protocol, namely the \textit{Istanbul} version defined in EIP-1679 by \cite{EIP-1679}.
In order to follow back the history of a path, one must reference multiple versions of this document.
Here are the block numbers of protocol updates on the Ethereum main network:
\par
\begin{center}
\begin{tabular}{lr}
\toprule
Name & First Block Number \\
\midrule
$F_{\mathrm{Homestead}}$         &  1150000 \\
$F_{\mathrm{Tangerine Whistle}}$ &  2463000 \\
$F_{\mathrm{Spurious Dragon}}$   &  2675000 \\
$F_{\mathrm{Byzantium}}$         &  4370000 \\
$F_{\mathrm{Constantinople}}$    &  7280000 \\
$F_{\mathrm{Petersburg}}$        &  7280000 \\
$F_{\mathrm{Istanbul}}$          &  9069000 \\
$F_{\mathrm{Muir Glacier}}$      &  9200000 \\
$F_{\mathrm{Berlin}}$            & 12244000 \\
\bottomrule
\end{tabular}
\end{center}
\par

Occasionally actors do not agree on a protocol change, and a permanent fork occurs.
In order to distinguish between diverged blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
For the Ethereum main network
\begin{equation}
  \linkdest{chain_id}
  \beta = 1
\end{equation}

@@

&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
\begin{cases}
\lVert \mathbf{b} \rVert & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] \neq \varnothing \\

@@

&&&& We assume $\mathbf{b} \equiv ()$ if $\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] = \varnothing$. \\

@@

@@

\begin{tabu}{\usetabu{opcodes}}
\midrule
0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current\\
&&&& environment. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
\midrule
0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
\midrule
\linkdest{extcodehash}{}0x3f & {\small EXTCODEHASH} & 1 & 1 & Get hash of an account's code. \\
&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
\begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}) \\ \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}} & \text{otherwise} \end{cases}$ \\
\bottomrule
\end{tabu}

@@

G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
